---
title: "A reciprocal history-conditioned pharmacology workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A reciprocal history-conditioned pharmacology workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(histPharm)
```

## Design principle

A reciprocal experiment contains matched `A_to_B` and `B_to_A` dose surfaces. Each biological replicate retains the vehicle, A-only margins, B-only margins, and combination cells. These controls allow the effect of the second drug after first-drug history to be compared with the same-phase second-drug effect without that history.

## Simulate known truth

```{r}
x <- simulate_sequence_data(
  h_A_to_B = 0.5,
  h_B_to_A = 0.1,
  n_models = 3,
  n_replicates = 3,
  seed = 11
)
x <- collapse_technical_replicates(x)
x <- normalize_sequence_data(x, baseline = "shared_replicate")
```

## Estimate interaction and absolute burden separately

```{r}
h <- estimate_history_effect(x)
dh <- estimate_reciprocal_asymmetry(x)
abs <- estimate_absolute_sequence_effect(x)

head(dh)
head(abs)
```

`delta_history` is not a synonym for absolute efficacy. The former asks whether the first-treatment history changes the second-drug incremental effect differently by order. The latter asks which matched schedule leaves less observed burden.

## Quantify uncertainty

```{r}
bootstrap_sequence_metrics(dh, "delta_history", B = 200, seed = 1)
```

For publication-scale analyses, use the prespecified biological replicates and increase the bootstrap draws.

## Transform to effect space

```{r}
e <- transform_effect_space(x)
head(e)
```

This separates nominal concentration from the biological perturbation depth actually produced by each single-drug margin.

## Benchmark an established sequential-response implementation

```{r eval=FALSE}
dchain_input <- as_dchain_viability_data(
  x,
  sequence = "A_to_B",
  first_dose = 1,
  drug_A = "DrugA",
  drug_B = "DrugB"
)
utils::write.table(
  dchain_input,
  "dchain_input.csv",
  sep = ",",
  row.names = FALSE,
  col.names = TRUE,
  quote = FALSE
)
```

The adapter reproduces the public example column schema. The public d-chain C++ parser expects an unquoted comma-separated file, so the example explicitly disables quoting. The adapter does not make histPharm and d-chain mathematically equivalent; the frameworks estimate different quantities and should be compared accordingly.
